---
title: "Homework 1"
author: "Harrison Wei"
output: github_document
---

## Problem 1

```{r penguins-data}
data("penguins", package = "palmerpenguins")
```

The `penguins` data have one row per penguin, with `r nrow(penguins)` rows and `r ncol(penguins)` columns. The species are `r paste(levels(penguins$species), collapse = ", ")`, and the islands are `r paste(levels(penguins$island), collapse = ", ")`. Other variables include bill length, flipper length, and body mass (all recorded as measurements), as well as sex and year. The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 1)` mm; this calculation excludes `r sum(is.na(penguins$flipper_length_mm))` missing values.

```{r penguins-plot}
library(ggplot2)

penguin_plot <- ggplot(
  penguins,
  aes(x = bill_length_mm, y = flipper_length_mm, color = species)
) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "Bill length (mm)",
    y = "Flipper length (mm)",
    color = "Species"
  ) +
  theme_minimal()

penguin_plot

ggsave(
  filename = "penguins_scatterplot.png",
  plot = penguin_plot,
  width = 7,
  height = 5,
  dpi = 300
)
```

The points suggest that Gentoo penguins tend to have longer flippers than the other two species. The scatterplot is also saved as `penguins_scatterplot.png` in the project directory.

## Problem 2
